---
title: "Lab 3 - Introduction to RStudio and R Markdown for reports and presentations"
output:
  pdf_document:
  toc: true
---
  
  <!-- This is a comment in R Markdown (and regular markdown, such as in Jupyter)-->
  <!-- The dashes below are not part of the YAML block above, -->
  <!-- they insert a horizontal line after the ToC -->
  
  --- 
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```
BEGIN ASSIGNMENT
requirements: requirements.R
generate: true
export_cell: false
files: 
  - imgs
```

In Lab 3,
you will learn how to use,
edit and create a R Markdown document (like this one) using RStudio.
You should follow the instructions in this document to complete the assignment.
Knit this document to view the nicely rendered HTML,
which can make it easier to read the questions.

If you need help as you use R Markdown in this lab and others in the future, consult the following resources:
  
- [Cheat sheet](https://rmarkdown.rstudio.com/lesson-15.html) 
- [Home page with guides](https://rmarkdown.rstudio.com/docs/)
- [Reference book](https://bookdown.org/yihui/rmarkdown/)

The below is a code chunk, 
but instead of using the `r` engine
we're creating and alert block
that will make the question show up with a blue background 
in the HTML output.
Unfortunately,
this creates and error when exporting to PDF,
so it can only be used for HTML.

```
BEGIN QUESTION
name: submission
manual: true
```

```{block, type="alert alert-info"}
## Submission Instructions
rubric={mechanics:2}
You receive mark for submitting your lab correctly, please follow these instructions:
- [Follow the general lab instructions](https://ubc-mds.github.io/resources_pages/general_lab_instructions/).
- [Click here to view a description of the rubrics used to grade the questions](https://github.com/UBC-MDS/public/tree/master/rubric).
- Push your `.Rmd` AND `.html` file to your GitHub repository for this lab.
    - The reason for pushing both files is that `.Rmd` does not contain the rendered output
      from running the cells.
      If someone is checking out your work
      there needs to be an HTML file to view the output,
      so it is good to get into this habit.
          - `.ipynb` renders nicely on GitHub,
            which is why we did not include the HTML file for previous labs.
- Upload a `.Rmd` version of your assignment to Canvas.
- Include a clickable link to your GitHub repo for the lab just below this cell (it should look something like this https://github.ubc.ca/MDS-2020-21/DSCI_521_labX_yourcwl.
```

<!-- BEGIN PROMPT -->

YOUR LAB GITHUB REPO LINK GOES HERE

<!-- END PROMPT -->

<!-- BEGIN SOLUTION NO PROMPT -->

> Solution: Correct link pasted
<!-- END SOLUTION -->

## Editing R Markdown documents

This document is called an R Markdown document.
It is a literate code document,
similar to Jupyter notebooks where you can write code and view its outputs.
To start,
let's set our working directory
by creating a new R Project for lab 3.

### Text and rendering R Markdown documents

In a R Markdown document any line of text not in a code chunk
(like this line of text)
will be formatted using Markdown.
Similar to JupyterLab,
you can also use HTML and LaTeX here to do more advanced formatting.
To run a code chunk,
you can press the green play button
in the top right corner of the chunk.

```
BEGIN QUESTION
name: q1
manual: true
```

```{block, type="alert alert-info"}
#### Question 1
rubric={correctness:1}
Since GitHub does not render HTML files by default,
you need to preface the raw URL to your rendered HTML files with the following 
in order to view them online: `http://htmlpreview.github.io/?`.
Using this approach, include a clickable link to your rendered HTML file on GitHub
under this question.
```

<!-- BEGIN PROMPT -->
  
  YOUR ANSWER GOES HERE

<!-- END PROMPT -->
  
  <!-- BEGIN SOLUTION NO PROMPT -->
  
  > Solution: Correct link pasted
<!-- END SOLUTION -->
  
  ```
BEGIN QUESTION
name: q2
manual: true
```

```{block, type="alert alert-info"}
#### Question 2
rubric={mechanics:1}
Create a new code chunk below using the r language engine that runs some R code
(it does not need to be complicated,
  but it should have an output).
Ensure that you can render/knit the document after you add that chunk.
```

<!-- BEGIN PROMPT -->
  
  YOUR ANSWER GOES HERE

<!-- END PROMPT -->
  
  <!-- BEGIN SOLUTION NO PROMPT -->
  
  > Solution: Any valid R code block with an output created.
<!-- END SOLUTION -->
  
  ```
BEGIN QUESTION
name: q3
manual: true
```

```{block, type="alert alert-info"}
#### Question 3
rubric={mechanics:1}
Create a new code chunk,
and add a meaningful name both to this and your previous code chunk.
Try using the pop-up-like menu to navigate between the named code chunks
Don't forget to knit/render the document
after you make this change to ensure everything is still working.
```

<!-- BEGIN PROMPT -->

YOUR ANSWER GOES HERE

<!-- END PROMPT -->

<!-- BEGIN SOLUTION NO PROMPT -->

> Solution: Note that students had a slightly different question here originally, so they might just have renamed the code block above which is fine for this year.
<!-- END SOLUTION -->

```
BEGIN QUESTION
name: q4
manual: true
```

```{block, type="alert alert-info"}
#### Question 4
rubric={mechanics:1,reasoning:1}
Create a new code chunk that uses a code chunk option.
Write out in your own words what the code chunk option is doing.
```

<!-- BEGIN PROMPT -->

YOUR ANSWER GOES HERE

<!-- END PROMPT -->

<!-- BEGIN SOLUTION NO PROMPT -->

> Solution: Any valid R code block with a chunk option such as `echo = FALSE`.
<!-- END SOLUTION -->

####  Multiple code chunk options

To have multiple code chunk options you separate them by a comma.
For example,
if in addition to suppressing warnings,
we want to run the code but not output the results,
then we can add the `include = FALSE` argument to the code chunk
after the `warning = FALSE` option.

```
BEGIN QUESTION
name: q5
manual: true
```

```{block, type="alert alert-info"}
#### Question 5
rubric={mechanics:1,reasoning:1}
Create a new code chunk that uses at least two code chunk options.
At least one must be different to the ones mentioned above.
Write in your own words what each code chunk option is doing.
```

<!-- BEGIN PROMPT -->

YOUR ANSWER GOES HERE

<!-- END PROMPT -->

<!-- BEGIN SOLUTION NO PROMPT -->

> Solution: Valid code chunk with two options and correct description.
<!-- END SOLUTION -->

### 1.5. YAML Header and document output options

R Markdown files contains three types of content:

1. Plain text mixed with simple Markdown formatting.
2. Code chunks surrounded by <code>```</code>.
3. An (optional) YAML header surrounded by `---`.
You have been introduced the first two types of content,
but not the third
(although you probably saw it at the top of this document).
The (optional) YAML header,
which is located at the very top of R Markdown files sets some general global parameters,
including:
- title
- author
- output
- etc
**Example YAML Header**
```
---
title: "Reproducible Data Science Report"
author: "Tiffany Timbers"
date: "September 4, 2018"
output: html_document
---
```
Most important from a workflow perspective is **output**. Possible output options include:
- `output: html_document`
- `output: md_document`
- `output: pdf_document`
- `output: word_document`
- `output: beamer_presentation` (beamer slideshow - pdf)
- `output: xaringan::moon_reader` (xaringan presentation - html)
```
BEGIN QUESTION
name: q6
manual: true
```
```{block, type="alert alert-info"}
#### Question 6
rubric={mechanics:1}
Navigate to the YAML header at the very top of this document
and edit it so that you include an `author` (yourself) and a `date` (lab due date).
Include what you added below here as well as a fenced Markdown code block.
```

<!-- BEGIN PROMPT -->

YOUR ANSWER GOES HERE

<!-- END PROMPT -->

<!-- BEGIN SOLUTION NO PROMPT -->

> Solution: It is ok if students only modified the header but did not add it here since the question was originally different. This might not show up in the html, so you will need to check the .Rmd in the repo.
<!-- END SOLUTION -->

```
BEGIN QUESTION
name: q7
manual: true
```

```{block, type="alert alert-info"}
#### Question 7
rubric={mechanics:1}
**This question previously mentioned to export the current document (lab3.Rmd).
to pdf.
This will not work due to the blue highlight boxes,
so please follow these updated instructions instead.**
After you have completed Question 8 and pasted the link to the HTML output,
navigate to the YAML header at the very top of that .Rmd document
and edit it so that the `output` is `pdf_document`.
Then knit/render the document.
Note the different output.
Add and commit that rendered `.pdf` file
to the GitHub repository for this lab
and paste a link to that file below this question.
```

<!-- BEGIN PROMPT -->

YOUR ANSWER GOES HERE

<!-- END PROMPT -->

<!-- BEGIN SOLUTION NO PROMPT -->

> Solution: Correctly exported PDf in the repo and a link here.
<!-- END SOLUTION -->

### Creating R Markdown documents

You can use the "File" menu inside RStudio to create new R Markdown documents by selecting:
File > New File > R Markdown...
This will bring you to another menu where you can choose the type of output
(don't be afraid to pick something,
you can always change the `output` type once you have the `.Rmd` file).

To create a written report,
we generally recommend using the default `output: html_document`
as it is easier to read than PDF
(note - LaTeX does not render nicely in such documents sadly,
  so if you are using a lot of LaTeX then you may want to choose `output: pdf_document`).
If you want to create an `.md` file to publish on GitHub,
it is recommend to instead use `output: github_document`.
To get this from the menu above
you need to navigate to the "From Template" option on the left panel
and then select "GitHub Document (Markdown)".

To create a presentation we recommend using xaringan.
To get a xaringan presentation from the File > New File > R Markdown... menu,
you need to navigate to the "From Template" option on the left panel
and then select "Ninja Presentation".
I will let the package author explain why it is named that way:
  
  > [Sharingan is] an R package for creating slideshows with remark.js through R Markdown.
> The package name xaringan comes from Sharingan,
> a dojutsu in Naruto with two abilities:
  > the "Eye of Insight" and the "Eye of Hypnotism".
> A presentation ninja should have these basic abilities,
> and I think remark.js may help you acquire these abilities,
> even if you are not a member of the Uchiha clan.
>
  > https://github.com/yihui/xaringan
When you create R Markdown documents this way,
RStudio gives you a template
and some reasonable defaults to start with
to help you get going quickly and easily.

```
BEGIN QUESTION
name: q8
manual: true
```

```{block, type="alert alert-info"}
#### Question 8
rubric={mechanics:6}
1. Create a new RMarkdown report
(a different file than this one) in the same directory as this Markdown file.
Use `html_document` as the `output`.
2. Create at least two Markdown text sections
(each should have a header)
and at least two separate code chunks in it (these can be really simple).
Save the new R Markdown document and  give it a meaningful name.
3. Render/knit the new R Markdown document to get an `.html` file.
Put the `.Rmd` document
and the rendered `.html` file under version control using Git,
and push/upload the file to your GitHub repository for this homework.
Paste a link to these files as your answer below.
```

<!-- BEGIN PROMPT -->
  
  YOUR ANSWER GOES HERE

<!-- END PROMPT -->
  
  <!-- BEGIN SOLUTION NO PROMPT -->
  
  > Solution: Correctly exported html and Rmd in the repo and a link here.
<!-- END SOLUTION -->
  
  ```
BEGIN QUESTION
name: q9
manual: true
```

```{block, type="alert alert-info"}
#### Question 9 (Optional)
rubric={mechanics:1,reasoning:1}
1. Take the R Markdown report created in Question 8
and change the output to `github_document` and render it.
Put the rendered `.md` file under version control using Git,
and push/upload the file to your GitHub repository for this homework.
Try to look at the file on GitHub.ubc.ca in your homework repo?
  What do you see? How is it rendered?
  ```

<!-- BEGIN PROMPT -->
  
  YOUR ANSWER GOES HERE

<!-- END PROMPT -->
  
  <!-- BEGIN SOLUTION NO PROMPT -->
  
  > Solution: `.md` document in the repo. Describe that GitHub renders the markdown file.
<!-- END SOLUTION -->
  
  ```
BEGIN QUESTION
name: q10
manual: true
```

```{block, type="alert alert-info"}
#### Question 10
rubric={mechanics:6}
1. Create a presentation using R Markdown.
Do this in a different file than this one but in the same directory as this Markdown file.
Use xaringan (see [this tutorial](https://bookdown.org/yihui/rmarkdown/xaringan.html)) as the presentation type.
Give this file a meaningful name.
2. Create at least 4 slides.
At least two slides must include a code chunk (these can be really simple).
Save the new R Markdown document.
3. Render/knit the new R Markdown document to get a `html` presentation file.
4. Put the new R Markdown document and the rendered `.html` file under version control using Git,
and push/upload the file to your GitHub repository for this lab
5. Open the `.html` presentation file and print/export to PDF from your browser.
Put the PDF under version control using Git,
and push/upload the file to your GitHub repository for this homework
(this is so you have a file in your homework repo that we can preview).
5. Paste a link to these files (`.Rmd`, .`html`, and `.pdf`) as your answer below.
```

<!-- BEGIN PROMPT -->
  
  YOUR ANSWER GOES HERE

<!-- END PROMPT -->
  
  <!-- BEGIN SOLUTION NO PROMPT -->
  
  > Solution: Link to the files, and make sure the exported html and PDF looks reasonable.
<!-- END SOLUTION -->
  
  ```
BEGIN QUESTION
name: q11
manual: true
```

```{block, type="alert alert-info"}
#### (Optional) Question 11
rubric={reasoning:1}
In a paragraph or two,
compare and contrast the use of reproducible tools
(e.g., R Markdown and Jupyter)
and non-reproducible tools
(Word, Powerpoint, Keynote, etc)
for presentations and reports.
Include advantages and disadvantages for each.
```

<!-- BEGIN PROMPT -->
  
  YOUR ANSWER GOES HERE

<!-- END PROMPT -->
  
  <!-- BEGIN SOLUTION NO PROMPT -->
  
  > Solution: Reproducible tools often makes it easier to collaborate long term and to track changes made by different people as they faciliate version control, as well as makes it easier to share and have someone else recreate your work and pomotes consistent formatting across slides and presentations. Non-reproduciple tools often have a lower barrier of entry, especially as people usually have used similar tools before. They also allow for live collaboration which is sometimes needed, and to create more fine-grained manual edits to graphics, such as image placement and aniations in presentations.
<!-- END SOLUTION -->
  
  ```
BEGIN QUESTION
name: placeholder
manual: false
```

```{r}
## Hidden Test ##
testthat::expect_true(TRUE)
```